\documentclass[12pt]{amsart}
\input{aux/style}
\input{aux/usualcmds}
\addbibresource{../usualpapers.bib}
\addbibresource{../bibliography.bib}
\input{aux/commands}

\usepackage{csquotes}
\usepackage{enumitem}
\usepackage[letterpaper, margin=1.87cm]{geometry} % Sets paper size and margins
\setlength{\parindent}{0pt}
\usepackage{setspace} % For line spacing
\usepackage{mathptmx} % Times New Roman for both text and math
\usepackage[scaled=.92]{helvet} % Helvetica, scaled to match Times New Roman

% set up 12pt Times New Roman
\renewcommand{\rmdefault}{ptm}
\renewcommand{\sfdefault}{phv}
\renewcommand{\ttdefault}{pcr}

% line spacing
%\onehalfspacing % Alternatively, you can use \singlespacing to try to meet the "no more than six lines per inch" requirement

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5

\title[Effective algebro-homotopical constructions]{Effective algebro-homotopical constructions \\ with applications to data science and quantum field theory}

\author{Anibal~M.~Medina-Mardones}
%\address{A.M-M., Max Planck Institute for Mathematics \and University of Notre Dame}
%\email{\href{mailto:ammedmar@mpim-bonn.mpg.de}{ammedmar@mpim-bonn.mpg.de}}

\begin{document}
	\vspace*{-0pt}
	\maketitle
	\thispagestyle{empty}
	\input{sec/introduction}
	\input{sec/background}
	\input{sec/constructions}
	\input{sec/persistence}
	\input{sec/physics}
	\newpage
	\sloppy
	\printbibliography
\end{document}

Proposal

Five pages max.

Addressing the points below, describe the proposed research to be supported. Images and graphics are included in the page limit.

Recent progress

Describe your recent progress in research activities related to the proposal; for returning grantees, describe as well the progress attributable to your previous Discovery Grant.

Objectives

Define the short- and long-term objectives of your research program. Note that a research program should have a long-term vision that expands beyond the five years of the Discovery Grant. A single, short-term project or collection of projects does not constitute a research program.

Literature review

Discuss the literature pertinent to the proposal, placing the proposed research in the context of the state of the art.

Methodology

Describe the methods and proposed approach, providing sufficient details to allow the reviewers to assess the feasibility of the research activities.

Considering equity, diversity and inclusion (EDI) in the research process promotes research excellence by making research outcomes more ethically sound, rigorous, reproducible, and useful. It is important to consider EDI through each stage of the research process including, but not limited to, the research questions, design, methodology, analysis, interpretation and dissemination of results, and integrate these considerations where relevant. Consult Equity, diversity and inclusion considerations at each stage of the research process for more information.

Impact

Explain the anticipated significance of the work.

Note:

If the information provided is insufficient, NSERC reserves the right to take appropriate action, such as not soliciting reports from external reviewers or withdrawing the application from the competition.
If relevant to your research, consult NSERC’s Guidelines for the preparation and review of applications in interdisciplinary research and/or Guidelines for the preparation and review of applications in engineering and the applied sciences.
If NSERC determines that the subject matter is outside of its mandate at any time during the review cycle, the application will be rejected. See This link will take you to another Web site Selecting the appropriate federal granting agency for more information.
