% !TEX root = ../proposal.tex

\smallskip\textsc{Background}.
As mentioned in the introduction, Poincaré's definition of the cohomology of a space $X$ relies on its cochains, denoted simply as $\cochains$, which arise from a cellular decomposition of $X$, typically a triangulation.
This construction is local in nature, allowing for a piecewise analysis of $X$, and is thus well-suited for concrete computations.
The computer algebra system (CAS) \texttt{SageMath}, among others, is a tool where this approach to cohomology is implemented.
Unfortunately, cohomology—or equivalently, the quasi-isomorphism type of $\cochains$—omits crucial homotopical information.
As an illustrative example, consider the spaces $\bC P^2$ and $S^2 \vee S^4$, the union of a 2- and a 4-sphere over a point.
Despite their cohomology groups being isomorphic, these spaces are not homotopy equivalent.
To distinguish them, one can employ the Alexander--Whitney product in $\cochains$ to reveal that their ring structures in cohomology are not isomorphic \cite{alexander1936ring, whitney1938products}.
This product-enhanced version of $\cochains$ retains the original's computability and is also supported in \texttt{SageMath}.
Considering the suspension of these spaces leads to two non-homotopic spaces $\Sigma(\mathbb{C} P^2)$ and $\Sigma(S^2 \vee S^4)$ whose cohomologies are isomorphic as graded rings.
To distinguish them we can consider Steenrod squares on $\rH^\bullet$ which are effectively computable through the cup-$i$ products \cite{steenrod1947products}, a structure on $\cochains$ also implemented in \texttt{SageMath}, which generalizes the Alexander-Whitney structure.
These enhancements on $\cochains$ would not suffice to distinguish a similar comparison involving a quaternionic projective space.
In that case, we can use odd prime power operations, but these are not available on mainstream computer algebra systems yet.

A ground-breaking result in modern homotopy theory, due to Mandell \cite{mandell2001padic,mandell2006homotopy_type}, asserts that under certain finiteness assumptions, all the homotopical information of $X$ is captured by an enhancement of the cup-$i$ product structure on $\cochains$.
This complete invariant is known as an $E_\infty$-algebra and serves as the conceptual cornerstone of our program.
The author and others provided a concrete local description of this structure on $\cochains$ \cite{medina2020prop1, mcclure2003multivariable, berger2004combinatorial}, which opened the door for the development of effective constructions of progressively richer homotopical invariants.
Consequently, it enabled their potential integration into computer algebra systems, as well as into more concrete fields such as knot theory, toric geometry, and lattice quantum field theory.

%%As mentioned in the introduction, Poincar\'e's definition of the cohomology of a space $X$ relies on the cochains $\cochains$ of a cellular decomposition of $X$.
%%This construction is local and well-suited for concrete computations.
%%Unfortunately, cohomology, or equivalently the quasi-isomorphism type of $\cochains$, forgets much homotopical information.
%%For example, the cohomology groups of $\mathbb{C} P^2$ and $S^2 \vee S^4$, the union of a $2$- and a $4$-sphere over a point, are isomorphic despite these spaces not being homotopy equivalent.
%%By choosing triangulations for these spaces, the Alexander--Whitney product can be used to show that the ring structure in cohomology distinguishes them.
%%An algorithm for this computation is available for example on \texttt{SageMath}.
%
%
%A fundamental step in the direction of our goal is the description of a ring structure on $\rH^\bullet$ using the Alexander--Whitney product on simplicial cochains $\cochains$ \cite{alexander1936ring, whitney1938products}.
%Let us illustrate with an example the additional information captured by this ring structure.
%The cohomology groups of $\mathbb{C} P^2$ and $S^2 \vee S^4$, the union of a $2$- and a $4$-sphere over a point, are isomorphic despite these spaces not being homotopy equivalent.
%By choosing triangulations for these spaces, the Alexander--Whitney product can be used to show that the ring structure in cohomology distinguishes them.
%An algorithm for this type of computation is available for example on \texttt{SAGE}.
%Considering the suspension of these spaces leads to two non-homotopic spaces $\Sigma(\mathbb{C} P^2)$ and $\Sigma(S^2 \vee S^4)$ whose cohomologies are isomorphic as graded rings.
%Distinguishing these effectively lead us to a novel step in the direction of our goal.
%Let us consider a prime $p$.
%Steenrod made $\rH^\bullet$ it into a module over his algebra $\cA_p$ \cite{steenrod1962cohomology}.
%This structure is explicitly induced from the $E_\infty$-algebra structure on simplicial cochains by means of the cup-$(p,i)$ products.
%These generalize to odd primes the cup-$i$ products of Steenrod, and were introduced and implemented, using effective versions of May's operadic methods \cite{may1970general}, in \cite{medina2021may_st,medina2021comch}.
%The action of $\cA_2$ suffices to distinguish $\Sigma(\mathbb{C} P^2)$ and $\Sigma(S^2 \vee S^4)$, but and odd prime will be needed for similar comparisons using the suspension of quaternionic projective spaces, for example.
%
%
%A ground-breaking result in modern homotopy theory, due to Mandell \cite{mandell2006homotopy_type}, is that, under certain finiteness assumptions, all the homotopical information of $X$ is captured by a so-called $E_\infty$-algebra structure on $\cochains$.
%This complete algebro-homotopical invariant is the conceptual cornerstone of our program, since, after work by the author and others, it can be given a concrete local description (\cite{mcclure2003multivariable,berger2004combinatorial,medina2020prop1}).
%This allows for the effective description of progressively richer homotopical invariants that can be implemented into computer algebra systems (CAS) such as \texttt{ComCH}, \texttt{SageMath}, and \texttt{Maple}.