\section{Current work} \label{s:current}

%\subsection{Multisimplicial chains and configuration spaces}
%
%Following a suggestion of P.~May, in joint work with A.~Pizzi and P.~Salvatore, the model discussed in \cref{ss:e-infty operads} is used to define a natural $E_\infty$-structure on multisimplicial chains.
%This is used to construct a zig-zag of quasi-isomorphisms of $E_\infty$-coalgebras between the singular chains of euclidean configuration spaces and the multisimplicial models of McClure--Smith.

\subsection{Framed polytopes, higher categories and Steenrod diagonals} \label{ss:polytopes}

In joint work with G. Laplante-Anfossi and B. Vallette an exploration of the convex geometry of higher diagonals of polytopes is presented.
The uncovered structures create a bridge between Steenrod type constructions and higher category theory.
More specifically, it is shown that both; a higher diagonal on a $n$-dimensional polytope $P$ and a free $n$-category structures on it can be canonically constructed from a generic frame of $P$.
The techniques developed are used to provide a full proof of a theorem of Kapranov and Voevodsky announced in \cite{kapranov1991polycategories}.

\subsection{Odd prime Steenrod operations: persistence theory and Khovanov homology} \label{ss:odd prime steenrod operations}

In joint work with F.~Cantero-Mor\'an, \emph{stable} cup-$(p,i)$ products of cochains are introduced.
There are three sources of motivation for this construction.
The first is to develop chain approximations to generalized homology theories extending those defined by Brumfiel--Morgan and discussed in more detail in \cref{ss:spt phases}.
The second is to generalize the constructions of Steenrod squares on Khovanov homology presented Cantero-Mor\'an in \cite{cantero-moran2020khovanov}.
The third is to obtain faster algorithms to compute Steenrod operations of finite simplicial complexes, facilitating the potential incorporation of this additional structure into topological data analysis as in \cref{ss:steenroder}.

\subsection{Geometric cohomology} \label{ss:geometric cohomology}

de Rham cohomology has long been lauded as a perfect cohomology theory by champions such as Sullivan and Bott.
A combination of geometric underpinning and commutativity at the cochain level make it a remarkably effective tool for many applications of rational homotopy theory.
Over the integers, submanifolds and intersection in various settings provide geometrically meaningful cochains with a partially defined commutative product.
Joint work with D. Sinha and G. Friedman provides a firm foundation for a model of ordinary (co)homology of smooth manifolds based on maps from manifolds with corners as generating (co)chains.
The first application of this theory already appeared in \cite{medina2021flowing} and was summarized above in \cref{ss:flowing}.
Several more applications of this theory are to follow.
Eventually, it is expected to geometrically encode the $E_\infty$-structure of manifolds in a geometric form.