% !TEX root = ../proposal.tex

\begin{tikzpicture}[remember picture, overlay]
	\node [anchor=north east, inner sep=0.5cm] at (current page.north east)
	{\large Anibal M. Medina-Mardones};
\end{tikzpicture}
\vspace*{-9pt}

\textsc{\large Effective Algebro-Homotopical Constructions with Applications to Data Science and Quantum Field Theory}
--- \hfill \textsc{\large Anibal M. Medina-Mardones}
\hrule

\bigskip\medskip
{\centering \underline{\textsc{Introduction}}\par}

\smallskip There is a tense trade-off in algebraic topology having roots reaching back to the beginning of its modern form.
This tension can be illustrated with the concept of cohomology, simply denoted $\coho$.
The first approaches, dating back to Poincar\'e, are based on the subdivision of a space into simple pieces.
% m
These elementary shapes and their spatial relations are made to generate a cochain complex from which $\coho$ is extracted.
This definition makes certain geometric properties of $\coho$, for example excision, fairly clear.
% m such as
Yet, it is not easy to show that a continuous map of spaces induces a map between their associated cohomologies.
The functoriality just alluded to is trivial when defining $\coho$ in terms of homotopy classes of maps to Eilenberg--MacLane spaces, but the passage to the homotopy category erases geometric and combinatorial information and the resulting definition is not well suited for concretely presented spaces.
The tension this example illustrates manifests itself in many other important contexts, and the trade-off between constructiveness and functoriality is as central today as it was almost a century ago.

\smallskip
{\centering\underline{\textsc{Objectives}}\par}
\smallskip\textit{My long-term vision is to ease this tension by developing concrete constructions of homotopical invariants currently defined only indirectly; and, by doing so, extend the reach of algebraic topology into fields beyond pure mathematics, including Data Science and Quantum Field Theory.}

\smallskip We will organize the presentation of the first five years of this program around four interrelated objectives.

\smallskip
(1) Compute primary \& secondary cohomology operations of triangulated spaces. \hfill (MSc1-2, PhD1-2)\par
(2) Compute Steenrod operations on Khovanov homology of knots and links. \hfill (MSc3, PhD3)\par
(3) Compute Steenrod barcodes and persistent cup-lengths of real-world datasets. \hfill (MSc4-5, PhD4)\par
(4) Build a bridge between state sum and functorial topological quantum field theories. \hfill (MSc6, PostDoc)\par